---
title: "Project 1"
author: "Joseph Kremor"
date: "Friday, February 02, 2018"
output: html_document
---
## Loading and preprocessing the data

```{r, echo=TRUE}


library(data.table)
library(lattice)

Data = as.data.table(read.csv('activity.csv'))

Data = Data[,date:=as.POSIXct(as.character(date))]

```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
Total_steps = Data[,mean(steps,na.rm = T),by = date]
# 
# make a histogram# 
hist(Total_steps$V1,main = 'Total Number of Steps per Day',xlab = 'Steps')


print(paste0('Daily mean: ',mean(Total_steps$V1,na.rm=TRUE)))
print(paste0('Daily median: ',median(Total_steps$V1,na.rm=TRUE)))

```

## What is the average daily activity pattern?
```{r, echo=TRUE}
Total_steps = Data[,mean(steps,na.rm = T),by = interval]
# 
# make a histogram# 
plot(Total_steps$interval,Total_steps$V1,main = 'daily step patterns',xlab = 'Five Minute Interval',ylab = 'Average Steps')


print(paste0('Interval with the most steps: ',Total_steps$interval[which.max(Total_steps$V1)]))

```

## Imputing missing values
```{r, echo=TRUE}
# Total number of missing values
print(paste0('Number of missing data: ',sum(!complete.cases(Data))))

Data_equal = copy(Data)

# Fill in missing values#
Data_equal$steps[which(is.na(Data_equal$steps))] <- mean(Data_equal$steps,na.rm=T)

Total_steps = Data_equal[,mean(steps,na.rm = T),by = date]

hist(Total_steps$V1,main = 'Total Number of Steps per Day Imputed',xlab = 'Steps')
# 
print(paste0('Daily mean: ',mean(Total_steps$V1,na.rm=TRUE)))
print(paste0('Daily median: ',median(Total_steps$V1,na.rm=TRUE)))

```

The effect of imputing missing data is that the variance of the distribution is decreased. The mean has not changed but the median has become closer to the mean.


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
# Weekends_factor
Data = Data[,weekends:='weekday']
Data = Data[which(weekdays(date)%in%c('Sunday','Saturday')),weekends:='weekend']
Data = Data[,weekends:=as.factor(weekends)]

Total_steps = Data[,mean(steps,na.rm = T),by = list(interval,weekends)]

xyplot(V1 ~ interval | factor(weekends), data=Total_steps,ylab = 'Steps',type = 'l',layout = c(1,2))

```

















